---
title: "Worksheet: EDA Worksheet"
subtitle: "STT 2860 Intro to Data Mgmt & Visualization" 
author: "Camden Lauver"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 4 
---

<hr>

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

## How We’ll Work Today (Groups + Individual Submission)

* **Use the same groups as Project 1.**
* You are encouraged to **talk, compare code, and help each other debug**, but **each student must submit their own completed worksheet to GitHub** (your own repo).
* If you get stuck, follow this order:

  1. read the error carefully
  2. ask your group
  3. ask me

---

## GitHub Workflow Reminder (Do This All Class)

**Before you start**

* Pull the most recent version (if needed).
* Open the `.Rmd`, knit once to make sure it runs.

**During work**

* Work in small steps.
* After each checkpoint section: **Knit → Commit → Push**

---

## Learning Notes (You will write as you go)

Create short notes while you work. You’ll fill these in at the end.

* **One thing I learned today:**
* **One thing that confused me at first but makes sense now:**
* **One ggplot trick I want to remember:**

---

# Packages Used

```{r packages}
# load all packages used here
library(tidyverse)
library(nycflights13)
library(skimr)
```

---

# Datasets Used

## Popular Kids

The dataset `popular.csv` and you can read about it [HERE](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/popular.html).

```{r kids-read}
# import the dataset from the web URL
kids <- read_csv("https://aokine.github.io/Data/popular.csv")
```

```{r kids-explore}
# explore the dataset
glimpse(kids)
skim(kids)
```



## Semmelweis Data

The dataset `semmelweisannualbyclinic.csv` and you can read about Dr. Ignaz Semmelweis [HERE](https://www.pbs.org/newshour/health/ignaz-semmelweis-doctor-prescribed-hand-washing). This dataset contains annual observations for two different clinics that led him to implement his handwashing experiment.

```{r semm-read}
semm <- read_csv("https://aokine.github.io/Data/semmelweisannualbyclinic.csv")
```

```{r semm-explore}
# explore the dataset
str(semm)
glimpse(semm)
skim(semm)
```

---

## Checkpoint 1 (Setup + Datasets)

**Checkpoint task:** Knit your document. If it works, **commit and push** with:

> `worksheet-eda: checkpoint-1 (setup + datasets)`

---

# NFL Players 2016 Dataset

Description: [https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/NFLplayers2016.html](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/NFLplayers2016.html)

## Load the dataset

```{r nfl-read}
NFL <- read_csv("https://aokine.github.io/Data/NFLplayers2016.csv")
```

## Explore the structure with `skim()`

```{r nfl-skim}
# Use skim()
```

---

# Barplots with `geom_bar()`

For `geom_bar()`, we input an `x` variable and the geometry counts how many of each value occurs in that variable, and that becomes the "y" value for the height of the bars.

## Barplot: Position Groups

Use `x = PositionGrp`.

```{r bar-1}
# TODO: Create barplot of PositionGrp
```

### Choose-your-own mini task (pick ONE)

1. Change bar color and outline (`fill =`, `color =`) **OR**
2. Flip coordinates (`coord_flip()`) **OR**
3. Make the bars show proportions instead of counts (hint: `aes(y = after_stat(count/sum(count)))`)

```{r bar-1-choice}
# Do ONE of the three options above. Add your code here.

```

---

## Fixing category names: Map `PositionGrp` to the y aesthetic

Make `PositionGrp` go on the **y-axis** so labels are easier to read.

```{r bar-2}
# TODO: Make horizontal barplot
```

---

## Barplot for Two Variables

Use `PositionGrp` and `Side` as the two variables. Use color/fill to separate groups.

```{r bar-3}
# TODO: Add fill = Side
```

### Quick interpretation (1–2 sentences)

In plain language, what does this plot suggest about how offense/defense/special teams are distributed across position groups?

**Answer:**

---

## Checkpoint 2 (Skim + geom_bar)

**Checkpoint task:** Knit your document. If it works, **commit and push** with:

> `worksheet-eda: checkpoint-2 (skim + first barplots)`

---

# Barplots with `geom_col()`

For `geom_col()`, we input an `x` variable and we input a `y` variable that specifies bar heights. That means we usually **summarize first**.

## Count the Position Groups

```{r col-1}
# TODO: Create summarized dataset using count()
```

## Plot counts with `geom_col()`

```{r col-2}
# TODO: Plot using geom_col()
```

---

## Rearrange the bars in order using `reorder()`

Change the order of the categories according to their counts (largest to smallest).

```{r col-3}
# TODO: Use reorder() to sort bars
```

### Choose-your-own mini task (pick ONE)

1. Add data labels above the bars (hint: `geom_text(aes(label = n), ...)`) **OR**
2. Change to descending order clearly (you might need `-n`) **OR**
3. Show proportions instead of counts (create a `prop = n/sum(n)` column)

```{r col-choice}
# Do ONE of the three options above. Add your code here.

```

---

## Checkpoint 3 (Two-variable bars + geom_col)

**Checkpoint task:** Knit your document. If it works, **commit and push** with:

> `worksheet-eda: checkpoint-3 (two-variable bars + geom_col)`

---

# Facets

## Facet Wrap

### Basic Histogram using `Weight`

```{r hist-1}
# TODO: Basic histogram
```

### Histogram with Fill using `PositionGrp` OR `Side`

Choose ONE: `fill = PositionGrp` or `fill = Side`

```{r hist-2}
# TODO: Add fill
```

### Histogram using `facet_wrap()` with `PositionGrp`

```{r hist-3}
# TODO: facet_wrap by PositionGrp
```

### Histogram using `facet_grid()` with `Side`

```{r hist-4}
# TODO: facet_grid by Side
```

### Quick interpretation (2–3 sentences)

Pick one faceted plot and write what pattern you notice.

**Answer:**

---

## Checkpoint 4 (Reorder + facets)

**Checkpoint task:** Knit your document. If it works, **commit and push** with:

> `worksheet-eda: checkpoint-4 (reorder + facets)`

---

# Reflection (Individual)

Fill these in before you submit.

1. **One thing I learned today:**

2. **One thing that confused me at first but makes sense now:**

3. **One ggplot trick I want to remember:**

4. **Choose-your-own reflection (pick ONE):**

* A. One plot I’m proud of and why
* B. One plot I would redesign and how
* C. One dataset question I’d investigate next

**Answer:**

---

## Final Submission

**Final task:** Knit one last time. Then **commit and push** with:

> `worksheet-eda: final (reflection + cleanup)`

---

---

```{r}
sessionInfo()
```
